---
title: "Premier League Score Prediction"
author: "Andy White"
date: "29 August 2016"
output: html_document
---

### Season 2015/16

Load Premier League data from 2015/16 season

```{r}
library(data.table)
library(ggplot2)
pl_dt = fread("Premier League 201516.csv")
x = model.matrix(~ HomeTeam + AwayTeam, data=subset(pl_dt, select=c("HomeTeam", "AwayTeam")))[,-1]
y = pl_dt[,FTHG]
xstar = pl_dt[,HST]
ggplot(pl_dt) + geom_boxplot(aes(factor(FTHG), HST)) +
  xlab("Goals Scored by Home Team") +
  ylab("Shots on Target by Home Team") +
  ggtitle("Relationship between shots on target and goals scored")
```

Build regression model based just on team data - we use a poisson model here as goals are count data:

```{r}
x_model = glm(y ~ ., data=cbind.data.frame(x, y=y), family=poisson())
x_preds = predict(x_model, as.data.frame(x), type="response")
```

Now build a regression model based on team data and shots on goal. We use a ridge model here because team and shots on goal are collinear:

```{r}
library(glmnet)
xstar_model = cv.glmnet(cbind(x, xstar), y, family="poisson", nfolds=3)
xstar_preds = predict(xstar_model, cbind(x, xstar), s=xstar_model$lambda.min, type="response")
num_train = 75
```

The MSE for the x-based model is `r mean((x_preds - y)^2)`.

The MSE for the x*-based model is `r mean((xstar_preds - y)^2)`.

Now build a model on the first `r num_train` matches of the season, and test it on the last `r 380 - num_train` matches of the season:

```{r}
x_model = glm(y ~ ., data=cbind.data.frame(x, y=y)[1:num_train,], family=poisson())
x_preds = predict(x_model, as.data.frame(x)[(num_train+1):380,], type="response")
xstar_model = cv.glmnet(cbind(x, xstar)[1:num_train,], y[1:num_train], family="poisson", nfolds=3)
xstar_preds = predict(xstar_model, cbind(x, xstar)[(num_train+1):380,], s=xstar_model$lambda.min, type="response")
```

The out-of-sample MSE for the x-based model is `r mean((x_preds - y[(num_train+1):380])^2)`.

The out-of-sample MSE for the x*-based model is `r mean((xstar_preds - y[(num_train+1):380])^2)`.

Now build a LUPI model:

```{r}
s = predict(xstar_model, cbind(x, xstar)[1:num_train,], s=xstar_model$lambda.min, type="response")
lupi_model = glm(s ~ ., data=cbind.data.frame(x[1:num_train,], s=s[,1]), family=Gamma(link="log"))
lupi_preds = predict(lupi_model, as.data.frame(x)[(num_train+1):380,], type="response")
```

The out-of-sample MSE for the LUPI model is `r mean((lupi_preds - y[(num_train+1):380])^2)`.

Plot the results:

```{r}
# Unprivileged plot
ggplot(pl_dt[(num_train+1):380]) + geom_boxplot(aes(factor(FTHG), x_preds)) +
  xlab("Goals Scored by Home Team") +
  ylab("Prediction with Unprivileged Data") +
  ggtitle("Predicting Goals Scored Using Unprivileged Data")
# Privileged plot
ggplot(pl_dt[(num_train+1):380]) + geom_boxplot(aes(factor(FTHG), lupi_preds)) +
  xlab("Goals Scored by Home Team") +
  ylab("Prediction with LUPI") +
  ggtitle("Predicting Goals Scored Using LUPI")
```

### Aggregating for each season

```{r}
run_football_models = function(filename, num_train) {
  require(glmnet)
  pl_dt = fread(paste0("Premier League 20", filename, ".csv"))
  x = model.matrix(~ HomeTeam + AwayTeam, data=subset(pl_dt, select=c("HomeTeam", "AwayTeam")))[,-1]
  y = pl_dt[,FTHG]
  xstar = pl_dt[,HST]
  x_model = glm(y ~ ., data=cbind.data.frame(x, y=y)[1:num_train,], family=poisson())
  x_preds = predict(x_model, as.data.frame(x)[(num_train+1):380,], type="response")
  xstar_model = cv.glmnet(cbind(x, xstar)[1:num_train,], y[1:num_train], family="poisson", nfolds=3)
  xstar_preds = predict(xstar_model, cbind(x, xstar)[(num_train+1):380,], s=xstar_model$lambda.min, type="response")
  s = predict(xstar_model, cbind(x, xstar)[1:num_train,], s=xstar_model$lambda.min, type="response")
  lupi_model = glm(s ~ ., data=cbind.data.frame(x[1:num_train,], s=s[,1]), family=Gamma(link="log"))
  lupi_preds = predict(lupi_model, as.data.frame(x)[(num_train+1):380,], type="response")
  g_model = glm(xstar ~ ., data=cbind.data.frame(x, xstar=xstar)[1:num_train,], family=poisson())
  g_preds = predict(g_model, as.data.frame(x)[(num_train+1):380,], type="response")
  fstar_preds = predict(xstar_model, cbind(x[(num_train+1):380,], xstar=g_preds), s=xstar_model$lambda.min, type="response")
  x_mae = round(mean(abs(x_preds - y[(num_train+1):380])), 3)
  priv_mae = round(mean(abs(xstar_preds - y[(num_train+1):380])), 3)
  lupi_mae = round(mean(abs(lupi_preds - y[(num_train+1):380])), 3)
  naive_mae = round(mean(abs(fstar_preds - y[(num_train+1):380])), 3)
  return(list(`Unprivileged MAE`=x_mae, `Privileged MAE`=priv_mae, `Distilled LUPI MAE`=lupi_mae, `Naive LUPI MAE`=naive_mae))
}
filenames = c("0607", "0708", "0809", "0910", "1011", "1112", "1213", "1314", "1415", "1516")
test_results = do.call("rbind.data.frame", sapply(filenames, function(f) run_football_models(f, 75), simplify=F))
names(test_results) = c("Unprivileged", "Privileged", "Distilled LUPI", "Naive LUPI")
test_results$season = 2007:2016
library(reshape2)
ggplot(melt(test_results, id="season")) +
  geom_line(aes(x=season, y=value, colour=variable)) +
  geom_point(aes(x=season, y=value, colour=variable), alpha=0.5) +
  xlab("Season") + ylab("Mean Absolute Error (MAE)") + scale_colour_discrete(name="Model")
```

End.