---
title: "Xstar as linear combination of interactions"
author: "Andy White"
date: "17 August 2016"
output: html_document
---

Below we can see that a single interaction term relating $X$ to $X^*$ and another relating $X^*$ to $Y$ results in a highly unusual relationship between $X$ and $Y$. 

```{r}
library(ggplot2)
set.seed(1)
X = matrix(rnorm(2*10000), nrow=10000, ncol=2)
Xstar = cbind(X, X[,1] * X[,2]) %*% matrix(rnorm(3*2), nrow=3, ncol=2)
Y = cbind(Xstar, Xstar[,1] * Xstar[,2]) %*% matrix(rnorm(3), nrow=3, ncol=1)
Yclass = Y > median(Y)
ggplot() + geom_point(aes(Xstar[,1], Xstar[,2], colour=factor(Yclass))) + labs(x="X*1", y="X*2", main="Classification of data in X*")
ggplot() + geom_point(aes(X[,1], X[,2], colour=factor(Yclass))) + labs(x="X1", y="X2", main="Classification of data in X")
```

Below we have a 10-dimensional input set. $X^*$ is a random linear combination of $X$ including all 2-way interaction terms. This gives a total of `r sum(choose(10, 1:2))` input features.

$Y$ is a random linear combination of $X^*$ including all 2-way interaction terms. Again, this gives a total of `r sum(choose(10, 1:2))` input features.

```{r}
X = matrix(rnorm(10*10000), nrow=10000, ncol=10)
colnames(X) = paste0("x", 1:10)
interactions = combn(colnames(X), 2)
X_with_interactions = cbind(X, apply(interactions, 2, function(x) X[,x[1]] * X[,x[2]]))
Xstar = X_with_interactions %*% matrix(rnorm(ncol(X_with_interactions) * 10), nrow=ncol(X_with_interactions), ncol=10)
colnames(Xstar) = paste0("x", 1:10)
Xstar_with_interactions = cbind(Xstar, apply(interactions, 2, function(x) Xstar[,x[1]] * X[,x[2]]))
Y = Xstar_with_interactions %*% matrix(rnorm(ncol(Xstar_with_interactions)), nrow=ncol(Xstar_with_interactions), ncol=1)
Yclass = Y > median(Y)
ggplot() + geom_point(aes(Xstar[,1], Xstar[,2], colour=factor(Yclass))) + labs(x="X*1", y="X*2", main="Classification of data in X*")
ggplot() + geom_point(aes(X[,1], X[,2], colour=factor(Yclass))) + labs(x="X1", y="X2", main="Classification of data in X")
```

Now let's try and learn $Y$ from $X$ using all 4-way interaction terms and a logistic classifier:

```{r}
for (i in seq(100, 2000, 100)) {
  train_samples = sample(10000, i)
  
}
```

End.